In this section we explore briefly some experiments I have conducted, in order to understand if the methodology I had in mind in order to prove that $\ROSP\notin\P$ is a valuable approach.

To begin with, the rank-one matrix $S = \alpha \mathbf{p}\mathbf{p}^T$, with $\mathbf{p}\in\{-1,0,1\}^n$, $\alpha\in\mathbb{R}_{>0}$ is very general. 

In order to shrink the search space (especially $\alpha$), we might consider only a small subsets of $M$ that has a very strict structure, so that $\alpha = 1$ and $M$ can be interpreted as a combinatorial object rather than an algebraic one.

In my thughts, $M$ (symmetric) is:

\begin{equation}
\label{eq:matrix_M_combinatorial}
	M = \begin{bmatrix}
		n-2 & \pm 1 & \dots & \pm 1 \\
		\pm 1 & n - 2 & \dots & \pm 1 \\
		\vdots & \vdots & \ddots & \vdots \\
		\pm 1 & \pm 1 & \dots & n - 2
	\end{bmatrix}
\end{equation}

This is a good option due to the following reasons: 
\begin{itemize}
	\item $\mathbf{p}\in\{1,-1\}^n$. Since we need to adjust every row, no value in $\mathbf{p}$ can be zero.

	\item $\alpha=1$. This is because, since in every row we need to ``adjust'' at least half the elements, then it is convenient to push $\alpha$ to the largest value possible, which is 1. \hubi{We can seek a more formal argument for this but for now that will do.}
\end{itemize}

I will collect, in an unordered way, all the facts, intuitions and ideas I had about these matrices: why they look as a good candidate in order to show that $\ROSP\notin\P$ and what are the main difficulties to this aim.

\begin{intuition}
	To begin with, $M$ can be interepreted as a graph, with labels on the edges ($\pm 1$), and the task is: assign a color to every node, so that we ``guess right'' at least half + 1 edges. By ``guessing right'' we mean: if an edge has color $+$, then the two nodes need to have the same color; if the edge is colored $-$, then the colors on the edges need to be different. Then, this looks extremely similar to the Unique Game problem (at least as described on 
	\url{https://en.wikipedia.org/wiki/Unique_games_conjecture}, I am also reading \cite{svensson_approximation_2013}).
\end{intuition}

The main limitation to this view is that in the $\UG$ the fraction of constraints to satisfy is a global phenomena, in this setting we need to satisfy a certain fraction of constraints for every node.

Additionally, there is the issue (raised by Tullio the first time I talked to him and Koppany about it around September 2025) that the matrix $M$ of \Cref{eq:matrix_M_combinatorial} might not be PSD (and, hence, impossible to prove to be PSD using Gershgorin circles techniques \cite{deville_optimizing_2019,varga_gersgorin_2004}).

Luckily, this is not the case:

\begin{proposition}
\label{prop:combinatorial_M_is_psd}
	Matrices as in \Cref{eq:matrix_M_combinatorial} are almost always PSD (especially for large values of $n$), and when they are PSD, they can often be proved so using ROSP technique.
\end{proposition}

\begin{proof}
	The proof is only experimental, so it is not a proper proof: I created random symmetric matrices as in \Cref{eq:matrix_M_combinatorial}, checked that they are PSD and, indeed, they are.
	
	In addition to that, I also checked how many times you can prove that the matrix is PSD by using a shift vector $\mathbf{p}\in\{-1,1\}^n$: roughly half the times.
	
	To conclude, I checked if it is the case that the $\mathbf{p}$ vector can be something other than one of the rows: most of the times, it is just one of the rows (see \Cref{sec:tullio_approach_weighted_sum} for a further discussion on the topic). Sometimes, though ($\approx 10\%$) the vector is something different.
	
	In order to explore further the state of things, please have a look at the code in \texttt{check\_rosp.py} at commit \texttt{b770fbe} (or later?).
\end{proof}

The best problem I have at hand, so far, is:

\begin{question}
	Given a matrix $M$ as in \Cref{eq:matrix_M_combinatorial}---with the promise of being PSD---is it easy to find the right shift $S=\mathbf{pp}^T$? 
\end{question}

From a purely combinatorial point of view, the problems looks quite hard. Yet, although I would bet that it is not polynomially tractable, I would also imagine some weird property either algebraic or of the graph that canb be exploited in order to turn the problem into an easier one.

An additional observation to keep in mind, which might be useful in order to show the $\NP$-completeness of $\ROSP$: when assigning labels on the nodes, we are de-facto creating two cliques (the clique of $+1$'s and the clique of $-1$'s) with edges labelled as $+$, and a cut in between the two cliques with edges labelled as $-$.

This bears some similarities with $\MAXCUT$ (although, once again, $\MAXCUT$ is more of a general property of graphs, and here constraints are more local).

In addition, another consideration, is the number of edges labelled as $-$ is generally smaller than the number of edges with a $+$ sign: given a bipartition $A,B$ s.t. $A\cup B=V$, such that $A$ is labelled with a $+$ and $B$ with a $-$ sign, it holds the following.

\begin{lemma}
\label{lemma:more_plus_than_minus_edges}
	The number of edges with a $-$ sign oscillates between $n-1$ and $\frac{n^2}{4}$ (respectively, when $a$ the size of $A$ is $1$ or $\frac{n}{2}$), whereas the number of edges with a $+$ sign is in $[\frac{n^2}{4}-\frac{n}{2}, {n \choose 2} - (n-1)]$ (respectively, when $a=\frac{n}{2}$ and $a=n-1$).
\end{lemma}
\begin{proof}
	Just count them. When $a= \frac{n}{2}$ then the bipartition is perfect. Inside the two cliques we have $2 {\frac{n}{2} \choose 2}$ edges of sign $+$ which is $2 \frac{\frac{n}{2}(\frac{n}{2}-1)}{2} = \frac{n^2}{4} - \frac{n}{2}$. The size of the crossing cut (number of edges with $-$ label) is simply ${n\choose 2}-|A|=\frac{n^2}{4}$.
	
	When, instead, $a=1$ then the number of edges of sign $+$ is ${n-1\choose 2} = \frac{(n-1)(n-2)}{2} = \frac{n^2-3n-2}{2}$, and the number of edges with sign $-$ is simply $n-1$.
	
	In between, when $n-1>a>1$, the number of edges varies between these two extremes.
\end{proof}

The above \Cref{lemma:more_plus_than_minus_edges} gives us a very rough upper bound on the number of edges with minus sign there can be. 

\begin{question}
	Does it? Try to understand if it is truly the case that when $M$ is made of all $-1$'s, then the above \Cref{lemma:more_plus_than_minus_edges} is telling us that we cannot ``fix'' enough entries for every row. \hubi{I think it was a good idea, but it might be wrong. I would just need to count them.}
\end{question}
\begin{proof}
	Fast forward: I did it on top of my head with $n=10$: in practice, you have 90 entries (the 10 diagonal entries don't count), we have $8$ written on the diagonals, after the shift it becomes 7, so we can make at most 3 mistakes on every row, 30 mistakes in total. If they are all -1's, and we pick the combination $|A|=|B|=5$, which maximizes the number of -1's to $\frac{n^2}{4}$, multiplied by 2 to get $\frac{n^2}{2}=50$, then we still make 40 mistakes which are too many.
\end{proof}

This upper bound, though, could be not so interesting because a matrix with a lot of -1's is probably not PSD: I have done once again done some experiments, and it turns out that the matrix can still be PSD even if it has a fraction of roughly $1/200$ entries set to +1, in a matrix of size $100\times 100$. See commit \texttt{7a0a776} for further details.